---
title: Understanding Game Behaviour Among Adolescents
author: "Team cuRiosity"
format: 
  revealjs:
    theme: simple
    self-contained: true
editor: visual
---

```{r directory, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
final_skill <- read_csv("final_skill.csv")
allSkills_demoData <- read_csv("allSkills_demoData.csv")
final_skill_player_categories <- read_csv("final_skill_player_categories.csv")
player_categories <- read_csv("player_categories.csv")
rating_vs_completed <- read_csv("rating vs completed.csv")
rating_over_time <- read_csv("rating_over_time.csv")
plotted_data <- read_csv("plotted_data.csv")
categorised_rating_vs_completed <- read_csv("categorised_rating_vs_completed.csv")
examplePlayer <- read_csv("examplePlayer.csv")
```

# How do players behave as they progress through the game?

::: footer
MACQUARIE UNIVERSITY
:::

## 

We present the results of our **exploratory research** that aims to:

::: incremental
-   characterise play style by:
    -   investigating variations between different demographics
    -   tracking changes in skill prioritization over the progression of the game
    -   understanding the perseverance of players on their number of attempts
-   player perception vs game progression
:::

::: footer
MACQUARIE UNIVERSITY
:::

# 3 Ps:

::: incremental
-   

    # Prioritisation

-   

    # Perseverance

-   

    # Perception
:::

::: footer
MACQUARIE UNIVERSITY
:::

## Background on the data used:

-   This data is from a video game called Elm Street Stories. This data compiles participant's data and uses it to reduce risk taking behaviours and lowering the spread of human immunodeficiency virus (HIV) amongst adolescents.

::: footer
MACQUARIE UNIVERSITY
:::

## Data used

::: incremental
-   Demographic information of the created avatars, i.e. age, gender, ethnicity (this is for avatars only)

-   Happiness rating how satisfied the players were from their epilogue endings

-   Skill level avatar's final skill level at the end of the game

-   Risk metric in a playthrough each bad decision leads to a strike and after 3 strikes you fail that scenario

-   Percentage complete progression in the game to 100% i.e. player starts in a bad state and ends with an optimal state
:::

::: footer
MACQUARIE UNIVERSITY
:::

# Category 1: Prioritisation

::: footer
MACQUARIE UNIVERSITY
:::

## How did players prioritise different skills?

```{r overall}
ggplot(
  data = final_skill_player_categories %>% 
    group_by(skill_type, avatar_id) %>%
    summarise(avg_max = mean(value)),
  mapping = aes(
    x = skill_type, 
    y = avg_max,
    fill = skill_type
  )
) + geom_col() +
  labs(
    title = "Overall final skill levels"
  )
```

::: footer
MACQUARIE UNIVERSITY
:::

## Avatar Gender

```{r gender}
ggplot(
  data = final_skill_player_categories %>% 
    group_by(skill_type, avatar_gender) %>%
    mutate(avg = mean(value)) %>%
    slice_head(n = 1),
  mapping = aes(
    x = skill_type, 
    y = avg,
    fill = skill_type    
  )
) + geom_col() + facet_wrap(~ avatar_gender) +
  labs(
    title = "Overall final skill levels - by chosen gender"
  )
```

::: footer
MACQUARIE UNIVERSITY
:::

## Avatar Ethnicity

```{r ethnicity}
ggplot(
  data = final_skill_player_categories %>% 
    group_by(skill_type, avatar_id) %>%
    mutate(avg_max = mean(value)) %>%
    slice_head(n = 1),
  mapping = aes(
    x = skill_type, 
    y = avg_max,
    fill = skill_type
  )
) + geom_col() + facet_wrap(~ avatar_id) +
  labs(
    title = "Overall final skill levels - by chosen ethnicity"
  )
```

::: footer
MACQUARIE UNIVERSITY
:::

## Risk Metric

```{r risk metric}
ggplot(
  data = final_skill_player_categories %>% 
    group_by(skill_type, category) %>%
    mutate(avg_max = mean(value)) %>%
    slice_head(n = 1),
  mapping = aes(
    x = skill_type, 
    y = avg_max,
    fill = skill_type
    
  )
) + geom_col() + facet_wrap(~ category) +
  labs(
    title = "Overall final skill levels - by risk metric"
  )
```

::: footer
MACQUARIE UNIVERSITY
:::

## How did players progress through their skills?

```{r skills_demog_overall}
allSkills_demoData %>%
  ggplot(aes(x = attempt, y = meanSkillLvl, group = skill, colour = skill)) +
  geom_point() + 
  geom_line() +
  xlab('Attempt') + 
  ylab('Mean skill level')
```

::: footer
MACQUARIE UNIVERSITY
:::

## Avatar Ethnicity

```{r skills_demog_ethnicity}
library(ggplot2)
allSkills_demoData %>% 
  ggplot(aes(x = attempt, y = meanSkillLvl, group = skill, colour = skill)) +
  facet_wrap(~ avatar_id) +
  geom_point() + 
  geom_line() +
  xlab('Attempt') + 
  ylab('Mean skill level')
```

::: footer
MACQUARIE UNIVERSITY
:::

## Avatar Gender

```{r skills_demog_gender}
allSkills_demoData %>% 
  ggplot(aes(x = attempt, y = meanSkillLvl, group = skill, colour = skill)) +
  facet_wrap(~ avatar_gender) +
  geom_point() + 
  geom_line() +
  xlab('Attempt') + 
  ylab('Mean skill level')
```

::: footer
MACQUARIE UNIVERSITY
:::

## Schools

```{r skills_demog_school}
allSkills_demoData %>% 
  ggplot(aes(x = attempt, y = meanSkillLvl, group = skill, colour = skill)) +
  facet_wrap(~ school) +
  geom_point() + 
  geom_line() +
  xlab('Attempt') + 
  ylab('Mean skill level')
```

::: footer
MACQUARIE UNIVERSITY
:::

# Category 2: Perseverance

::: footer
MACQUARIE UNIVERSITY
:::

## How did players prioritise different skills?

Player attempts

```{r attempts_overall}
ggplot(allSkills_demoData %>%
         group_by(player_id) %>%
         summarise(max = max(attempt)), mapping = aes(x = max)) +
  geom_histogram(binwidth = 5) +
  xlab('Maximum Attempt') + 
  ylab('Count')

```

::: footer
MACQUARIE UNIVERSITY
:::

## By Gender

```{r attempt gender}
ggplot(allSkills_demoData %>%
       group_by(player_id) %>%
       mutate(max = max(attempt)) %>% 
         slice_head(n = 1), mapping = aes(x = max, color = avatar_gender, fill = avatar_gender)) +
  facet_wrap(~ avatar_gender) +
  geom_histogram(binwidth = 5) +
  xlab('Maximum Attempt') + 
  ylab('Count')

```

::: footer
MACQUARIE UNIVERSITY
:::

## By Ethnicity

```{r attempt ethnicity}
ggplot(allSkills_demoData %>%
         group_by(player_id) %>%
         mutate(max = max(attempt)) %>% 
         slice_head(n = 1), mapping = aes(x = max, fill = avatar_id)) +
  facet_wrap(~ avatar_id) +
  geom_histogram(binwidth = 5) +
  xlab('Maximum Attempt') + 
  ylab('Count')

```

::: footer
MACQUARIE UNIVERSITY
:::

## By School

```{r attempt school}
ggplot(allSkills_demoData %>%
         group_by(player_id) %>%
         mutate(max = max(attempt)) %>% 
         slice_head(n = 1), mapping = aes(x = max, fill = school)) +
  facet_wrap(~ school) +
  geom_histogram(binwidth = 5) +
  xlab('Maximum Attempt') + 
  ylab('Count')
```

::: footer
MACQUARIE UNIVERSITY
:::

# Category 3: Perception

::: footer
MACQUARIE UNIVERSITY
:::

```{r example}
completion_rating_over_time <- replace_na(fill(arrange(bind_rows(examplePlayer %>% 
  filter(event_id == 703) %>%
  select(event_time_dbl, data_values) %>% ## Note: Proportion_complete was broken
  separate(data_values, sep = ",", into = c("json_redacted","percentage_completed")) %>%
  mutate(percentage_completed = as.numeric(percentage_completed)),

# Over time their rating of the outcome. 
examplePlayer %>%
  filter(event_id == 707) %>%
  select(event_time_dbl, data_values) %>%
  separate(data_values, sep = ",", into = c("old_rating", "new_rating"))# %>%
  ##mutate(new_rating = as.numeric(new_rating))
), event_time_dbl), percentage_completed, new_rating), list(new_rating = "0"))

ggplot(
  data = completion_rating_over_time %>% 
    select(event_time_dbl, percentage_completed, new_rating) %>% 
    rename(Satisfaction_Rating = new_rating) %>%
    rename(Game_Time = event_time_dbl),
  mapping = aes(
    x = Game_Time,
    y = percentage_completed,
    fill = Satisfaction_Rating,
    
  )
)  + geom_area() + 
  scale_fill_brewer(palette = "RdYlGn") + 
  ylim(0, 1) +
  labs(title = "Player 6486040's Satisfaction During Game Play")

```

## How did players prioritise different skills?

::: panel-tabset
## Linear Graph

```{r plot }
ggplot(
  data = rating_vs_completed %>% select(player_id,percentage_completed, new_rating) %>%
    mutate(test = (as.numeric(new_rating) / 4) - percentage_completed) %>%
    group_by(percentage_completed) %>%
    summarise(
      avg = mean(test)
    ),
  mapping = aes(
    x = percentage_completed,
    y = avg
  )
) + geom_point() + 
  geom_smooth(method = lm) + 
  geom_hline(yintercept = 0) +
  ylim(-1, 1) +
  labs(
    y = "Variation"
  )
```

::: footer
MACQUARIE UNIVERSITY
:::

## Summary

```{r lm }
linear_model <- lm(data = plotted_data)
summary(linear_model)
```

::: footer
MACQUARIE UNIVERSITY
:::
:::

## Ethnicity

```{r lm ethnicity}
ggplot(data = categorised_rating_vs_completed %>% select(player_id,percentage_completed, new_rating, avatar_gender) %>%
    mutate(test = (as.numeric(new_rating) / 4) - percentage_completed) %>%
    group_by(percentage_completed, avatar_gender) %>%
    summarise(
      avg = mean(test)
    ) %>% drop_na(),
  mapping = aes(
    x = percentage_completed,
    y = avg
  )
) +
  
  geom_point() + 
  geom_smooth(method = lm) + 
  geom_hline(yintercept = 0) +facet_wrap(~ avatar_gender) + 
  ylim(-1, 1) +
  labs(
    y = "Variation"
  )
```

::: footer
MACQUARIE UNIVERSITY
:::

## Gender

```{r lm gender}
ggplot(
  data = categorised_rating_vs_completed %>% select(player_id,percentage_completed, new_rating, avatar_id) %>%
    mutate(test = (as.numeric(new_rating) / 4) - percentage_completed) %>%
    group_by(percentage_completed, avatar_id) %>%
    summarise(
      avg = mean(test)
    ) %>% drop_na(),
  mapping = aes(
    x = percentage_completed,
    y = avg
  )
) +
  
  geom_point() + 
  geom_smooth(method = lm) + 
  geom_hline(yintercept = 0) +facet_wrap(~ avatar_id) + 
  ylim(-1, 1) +
  labs(
    y = "Variation"
  )

```

::: footer
MACQUARIE UNIVERSITY
:::

## 

```{r density}
ggplot(
  data = categorised_rating_vs_completed %>% select(player_id,percentage_completed, new_rating, category) %>%
    mutate(new_rating = as.numeric(new_rating)) %>% drop_na()
  ,
  
  mapping = aes(
    x = percentage_completed,
    y = new_rating
  )
) + facet_wrap(~ category) + 
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette=4, direction=1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme( legend.position='none') +
  labs( title = "Density of satisfaction vs game completion" ) +
  ylab("Satisfaction Rating")
```

::: footer
MACQUARIE UNIVERSITY
:::

## Risk Levels

```{r lm risk metric}
ggplot(
  data = categorised_rating_vs_completed %>% select(player_id,percentage_completed, new_rating, category) %>%
    filter(category == "1. Low Risk") %>% group_by(percentage_completed) %>%
    mutate(test = (as.numeric(new_rating) / 4) - percentage_completed) %>%
    summarise(
      avg = mean(test)
    ),
  mapping = aes(
    x = percentage_completed,
    y = avg,
  )
) + 
  geom_smooth(method = lm, color="blue", size = 2) +  
  geom_smooth(method = lm,
              data = categorised_rating_vs_completed %>% select(player_id,percentage_completed, new_rating, category) %>%
                filter(category == "3. High Risk") %>% group_by(percentage_completed) %>%
                mutate(test = (as.numeric(new_rating) / 4) - percentage_completed) %>%
                summarise(avg = mean(test)),
              color = "red",
              size = 2
  )+   
  geom_smooth(method = lm,
              data = categorised_rating_vs_completed %>% select(player_id,percentage_completed, new_rating, category) %>%
                filter(category == "2. Medium Risk") %>% group_by(percentage_completed) %>%
                mutate(test = (as.numeric(new_rating) / 4) - percentage_completed) %>%
                summarise(avg = mean(test)),
              color = "yellow",
              size = 2
  )+ 
  geom_hline(yintercept = 0) +
  labs(y = "Variation")
```

::: footer
MACQUARIE UNIVERSITY
:::

# Future Recommendations

-   Analyse survey results with the different play styles of players
-   Determine whether outlets like this game can be used as a suitable method of teaching adolescents about risk behaviours
-   What extent can these outlets can help in reducing the risk taking behaviour evident in the bad decisions and endings of the games?

# Conclusion

1.  Effect of demographics was consistent across prioritisation and perseverance - no significant effect
2.  There is a strong potential for using the characteristics of perception as significant predictor of identifying individuals who are more inclined towards risk taking behaviours.

::: footer
MACQUARIE UNIVERSITY
:::
